---
title: "Data Preparation"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )

library(tidyverse)
```

# Overview

This Rmarkdown document summarises the data preparatory steps for the production of the talk/paper contribution to the Opensky Network Symposium, November 2022.

The paper aims to expand on the initial implementation of gate-to-gate fuel burn estimation by investigating / detailing taxi-modes and associated mode durations (i.e. time-in-mode).
The latter are essential to improve the overestimation of fuel burn during taxi based on the assumptions of the LTO cycle.

In general, fuel burn for a taxiing flight can be determined by summing the fuel burn per taxi-phase.
The ICAO Engine Emission Databank (IEED) (and complementary research) provides fuel flow rates at specific thrust settings.
A mapping of the aircraft type to "representative engines" supports then the calculation of the fuel burn per phase:

Fuel burn (during taxi) = sum of fuel burn per taxi phase.    
Fuel burn per taxi phase = number of engines * fuel flow rate * duration of phase (time-in-mode)

$$F_i = \sum_p (n_{engine} * f_{ip} * t_{ip}) $$

# Organisation

airport use-case data from Opensky Network --> needs to be downloaded/extracted.

For a start - use test data prepared by Enrico for other project: https://zenodo.org/record/6379014#.Yzxi7-yA5jM.
Data is stored in a top-level folder (to make it available for other projects).

RProjects
-- __DATA
-- paper-2022-OSN-SurfaceFuelBurn

# Data Airport Infrastructure

TODO
- download airport context from OSM --> filter
- potential to use runways from AIP or ourairports

by hand define "areas" of interest

# Surface Trajectory

To get a bit of a feel for what we have ... consult milestones generated by Dean/Enrico.

```{r}
mst <- readr::read_csv2("./data/ESSA-20210926-mst.csv", col_types = cols(.default = "c"))
mst <- mst %>% 
  mutate(timestamp = lubridate::ymd_hms(timestamp)
         ,longitude = as.numeric(longitude)
         ,latitude  = as.numeric(latitude)
         )

events <- mst %>% distinct(event_type)
events
```

ok. That is a start. As we deal with surface movements, we can safely discard any crossing (for now)

```{r}
mst <- mst %>% 
  filter(! event_type %in% c("c10","c40","c100"))
#mst
```

```{r}
pick_flt <- mst %>% filter(callsign %in% c("AEE761","AFL2211")) %>% 
  arrange(callsign, timestamp)
pick_flt
```

```{r}
# coerce pick_flt to sf-POINTS
pick_sf <- pick_flt %>%
    sf::st_as_sf(coords = c("longitude","latitude"), crs = 4326, remove = FALSE)

# run query_osm_apt.R for ESSA = baseplot g
essa <- g +
  geom_sf(data = pick_sf, aes(color = callsign, shape = event_type), size = 4)
essa
```



```{r}
sfc <- arrow::read_feather("./data-raw/ESSA/ESSA_20210926.feather")
```


